Hydras and Cie \\ Variations on Kirby \& Paris' hydra battles and ordinal numbers in Coq

### 2021-09-10 doc/JFLA2022/paper.tex
draft paper

%%%% For Alectryon

@@

% use this if you have a long article and want to create an index
% \usepackage{makeidx}

% In order to save space or manage large tables or figures in a
% landcape-like text, you can use the rotating and pdflscape
% packages. Uncomment the desired from the below.
%
% \usepackage{rotating}
% \usepackage{pdflscape}

% Some of our commands for this guide.
%
\newcommand{\easychair}{\textsf{easychair}}
\newcommand{\miktex}{MiK{\TeX}}
\newcommand{\texniccenter}{{\TeX}nicCenter}
\newcommand{\makefile}{\texttt{Makefile}}
\newcommand{\latexeditor}{LEd}
\newcommand{\canonseq}[2]{\mbox{$\{#1\}(#2)$}}

\newtheorem{todo}{To do}

@@

%% Front Matter
%%
% Regular title as in the article class.
%
\title{Hydras and Cie \\
  Variations on Kirby \& Paris' hydra battles and other entertaining math in Coq}

@@

\begin{abstract}
  We present the evolution of our library \cite{HydraBattles}
since \cite{JFLA2018paper}
\end{abstract}

\setcounter{tocdepth}{2}
{\small
\tableofcontents}

@@

\section{New contents}
We are interested in a usual variant of hydra-battles, where the hydra grows $n$ copies at step $n$ of the game (see the examples presented in ~\cite{KP82, BauerHydra}).
In terms of ordinals, this variant is described as a (finite) sequence $s_n\,(n\geq i)$ of ordinals strictly below $\epsilon_0$.
\begin{itemize}
\item If $s_n=0$, then the sequence is empty,
\item if $s_n=\beta+1$, then $s_{n+1}=\beta$,
\item Otherwise, $s_n$ is a \emph{limit ordinal} $\lambda$; then $s_{n+1}$ is $\canonseq{\lambda}{n+1}$, \emph{i.e.} the
    $n+1$-th element of the \emph{canonical sequence} associated with $\lambda$~\cite{KS81}.
\end{itemize}

The length of such a sequence is $l-i$, where $s_l=0$. It depends both on the initial index $i$ and the initial term $\alpha=s_i$. Adapting to \emph{Coq} a few lemmas from~\cite{KS81}, we prove that this length is minored by a fast growing function.
More precisely, the length  of the sequence associated with
$\omega^\alpha$ and initial index $i\geq 2$ is greater or equal than
$F_{\omega^\alpha}(i-1)-(i-1)$, where $F_\alpha$ is a variant of the Wainer hierarchy of rapidly growing functions, defined and studied in~\cite{KS81}.
From this minoration, we infer that the function which associates to $i$ the length of the considered sequence (for a given $\alpha$) is not primitive recursive in general (our counter-example is for $\alpha=\omega^\omega$).

>>>>

% -------------------------------
%%%% For Alectryon

@@

%---------------------------- 
\newcommand{\canonseq}[2]{\mbox{$\{#1\}(#2)$}}
\newtheorem{todo}{To do}
\usepackage{amsfonts}

\newcommand{\easychair}{\textsf{easychair}}
\newcommand{\miktex}{MiK{\TeX}}
\newcommand{\texniccenter}{{\TeX}nicCenter}
\newcommand{\makefile}{\texttt{Makefile}}
\newcommand{\latexeditor}{LEd}

@@

%% Front Matter
%%
% Regular title as in the article class.
%
\title{Hydras and Cie \\
  Variations on Kirby \& Paris' hydra battles and ordinal numbers in Coq}

@@

\begin{abstract}
  \begin{todo}
  We present the current state and evolution of a  library \cite{HydraBattles}
since \cite{JFLA2018paper}
  \end{todo}
\end{abstract}

% \setcounter{tocdepth}{2}
% {\small
% \tableofcontents}

@@

\section{New contents}
We are interested in a usual variant of hydra-battles, where the hydra grows $n$ copies at step $n$ of the game (see the examples presented in ~\cite{KP82, BauerHydra}).
In terms of ordinals, this variant is described as a (finite) sequence $s_n\,(n\geq i)$ of ordinals strictly below $\epsilon_0$.
This sequence is parameterized by an initial index $i\in\mathbb{N}$ and an ordinal $\alpha<\epsilon_0$.

\begin{itemize}
  \item $s_i=\alpha$
\item If $s_n=0$, then the sequence is empty,
\item if $s_n=\beta+1$, then $s_{n+1}=\beta$,
\item Otherwise, $s_n$ is a \emph{limit ordinal} $\lambda$; then $s_{n+1}$ is $\canonseq{\lambda}{n+1}$, \emph{i.e.} the
    $n+1$-th element of the \emph{canonical sequence} associated with $\lambda$~\cite{KS81}.
\end{itemize}

The length of such a sequence is $l-i$, where $l$ is the first --- and unique --- natural number such that $s_l=0$.
For a given $\alpha$, this length is a function of 
the initial index $i$.

Adapting to \emph{Coq} a few lemmas from~\cite{KS81}, we prove that the length  of the sequence associated with
$\omega^\alpha$ and initial index $i\geq 2$ is greater or equal than
$F_{\omega^\alpha}(i-1)-(i-1)$, where $F_\alpha$ is a variant of the Wainer hierarchy of rapidly growing functions, defined and studied in~\cite{KS81}.

From this minoration, we infer that the function which associates to $i$ the length of the considered sequence (for a given $\alpha$) is not primitive recursive in general (our counter-example is for $\alpha=\omega^\omega$).
